Title Text

### 2014-04-04 intromoa.tex
changes for psi reduction

The central operation of MoA is the indexing function
\[
                    p \psi A \]
in which a vector of $n$ integers $p$ is used to select an item of 
the $n$-dimensional array $A$.
The operation is generalized to select a partition of $A$, 
so that if $q$ has only $k < n $ components
then
\[                    q \psi A \]
is an array of dimensionality $n-k$ and $q$ selects among the possible 
choices for the first $k$ axes. In MoA zero origin indexing is assumed.
For example, if $A$ is the $3$ by $5$ by $4$ array
\[ \left [
\begin{array}{lrrrr}
0 & 1 &  2 & 3  \\
4 & 5 & 6 & 7 \\
8 & 9 & 10 & 11  \\
12 & 13 & 14 & 15 \\
16 & 17 & 18 & 19 
\end{array} \right ] 
\left [ \begin{array}{rrrr}
20 & 21 & 22 & 23 \\
24 & 25 & 26 & 27 \\
28 & 29 & 30 & 31 \\
32 & 33 & 34 & 35 \\
36 & 37 & 38 & 39 
\end{array} \right ]
\left [ \begin{array}{rrrr}
40 & 41 & 42 & 43 \\
44 & 45 & 46 & 47 \\
48 & 49 & 50 & 51 \\
52 & 53 & 54 & 55 \\
56 & 57 & 58 & 59
\end{array} \right ] \]
then
\[
     <1> \psi A = \left [  \begin{array}{rrrr}
                   20 & 21 & 22 & 23 \\
                   24 & 25 & 26 & 27 \\
                   28 & 29 & 30 & 31 \\
                   32 & 33 & 34 & 35 \\
                   36 & 37 & 38 & 39 
                    \end{array} \right ] \]

@@

%\break
\begin{figure}[h]
\scriptsize
\begin{tabular}{|l|l|l|}
\hline\hline
{\bf Symbol} & {\bf Name}& {\bf Description} \\ \hline
$\delta$ & Dimensionality & Returns the number of dimensions of an array.\\ \hline
$\rho$ & Shape & Returns a vector of the upper bounds \\
& &or sizes of each dimension in an array. \\ \hline
$\iota \xi^n$  & Iota & When $n=0$(scalar),
returns a vector containing elements\\
& & $0$, to $\xi^0 - 1$. When $n = 1$(vector), returns an \\
& & array of indices  defined by the shape vector $\xi^1$ \\ \hline
$\psi$ & Psi & The main indexing function of the Psi Calculus \\
& & which defines all operations in MoA.   Returns a scalar \\
& & if a full index is provided, a sub-array otherwise. \\ \hline
\mbox{rav} & Ravel & vectorizes a multi-dimensional array based   \\
&&on an  array's layout($\gamma_{row},\gamma_{col},\gamma_{sparse}, ...$)\\ \hline
$\gamma$ & Gamma & Translates indices into offsets given a shape. \\ \hline
$\gamma^{'}$ & Gamma Inverse & Translates offsets into indices given a shape. \\ \hline
$\vec s \rshp \xi $ & Reshape& Changes the shape vector of an array,
 possibly affecting  \\
& &its dimensionality. Reshape depends on
layout($\gamma$). \\ \hline
$\pi \vec x$ & Pi & Returns a scalar and is equivalent to $\prod_{i=0}^{(\tau x)-1}x[i]$ \\ \hline
$\tau$& Tau & Returns the number of components in an array,($\tau \xi \equiv
\pi (\rho \xi) $) \\ \hline
$\xi_l \cat \xi_r $ & Catenate& Concatenates two arrays over their
primary axis. \\ \hline
$\xi_l f \xi_r $ & Point-wise  & A data parallel application of $f$ is performed \\
&Extension &  between all elements of the arrays.\\ \hline
$\sigma f \xi_r $ & Scalar Extension & $\sigma$ is used with
every component of $\xi_r$ in the data parallel \\
$\xi_l f \sigma$ & &application of $f$. \\ \hline
$\take$&Take& Returns a sub-array from the beginning or end of an array\\
& & based on its argument being positive or negative. \\ \hline
$\drop$ & Drop & The inverse of Take \\ \hline
$_{op} \mbox{red}$&Reduce& Reduce an array's dimension by one by applying \\
& & op over the primary axis of an array. \\ \hline
$\Phi$& Reverse& Reverses the components of an array. \\ \hline
$\Theta $& Rotate & Rotates, or shifts cyclically, components of an
array. \\ \hline
$\transpose$ & Transpose & Transposes the elements of an array based on \\
&&a given permutation vector \\ \hline
$\Omega$& Omega& Applies a unary or binary function to array argument(s) \\
&&given partitioning information. $\Omega$ is used to perform all operations \\&& (defined over the primary axis only) over all dimensions. \\ \hline
\end{tabular}
\caption{\label{moa}Summary of MoA Operations}
\end{figure}

@@

>>>>

The central operation of MoA is the indexing function
\[
                    p \psi A \]
in which a vector of $n$ integers $p$ is used to select an item of 
the $n$-dimensional array $A$.
The operation is generalized to select a partition of $A$, 
so that if $q$ has only $k < n $ components
then
\[                    q \psi A \]
is an array of dimensionality $n-k$ and $q$ selects among the possible 
choices for the first $k$ axes. In MoA zero origin indexing is assumed.
For example, if $A$ is the $3$ by $5$ by $4$ array
\small
\[ \left [
\begin{array}{lrrrr}
0 & 1 &  2 & 3  \\
4 & 5 & 6 & 7 \\
8 & 9 & 10 & 11  \\
12 & 13 & 14 & 15 \\
16 & 17 & 18 & 19 
\end{array} \right ] 
\left [ \begin{array}{rrrr}
20 & 21 & 22 & 23 \\
24 & 25 & 26 & 27 \\
28 & 29 & 30 & 31 \\
32 & 33 & 34 & 35 \\
36 & 37 & 38 & 39 
\end{array} \right ]
\left [ \begin{array}{rrrr}
40 & 41 & 42 & 43 \\
44 & 45 & 46 & 47 \\
48 & 49 & 50 & 51 \\
52 & 53 & 54 & 55 \\
56 & 57 & 58 & 59
\end{array} \right ] \]
\normalsize
then
\[
     <1> \psi A = \left [  \begin{array}{rrrr}
                   20 & 21 & 22 & 23 \\
                   24 & 25 & 26 & 27 \\
                   28 & 29 & 30 & 31 \\
                   32 & 33 & 34 & 35 \\
                   36 & 37 & 38 & 39 
                    \end{array} \right ] \]

@@

%\break
\begin{figure}[h]
%\scriptsize
\tiny
\begin{tabular}{|l|l|l|}
\hline\hline
{\bf Symbol} & {\bf Name}& {\bf Description} \\ \hline
$\delta$ & Dimensionality & Returns the number of dimensions of an array.\\ \hline
$\rho$ & Shape & Returns a vector of the upper bounds \\
& &or sizes of each dimension in an array. \\ \hline
$\iota \xi^n$  & Iota & When $n=0$(scalar),
returns a vector containing elements\\
& & $0$, to $\xi^0 - 1$. When $n = 1$(vector), returns an \\
& & array of indices  defined by the shape vector $\xi^1$ \\ \hline
$\psi$ & Psi & The main indexing function of the Psi Calculus \\
& & which defines all operations in MoA.   Returns a scalar \\
& & if a full index is provided, a sub-array otherwise. \\ \hline
\mbox{rav} & Ravel & vectorizes a multi-dimensional array based   \\
&&on an  array's layout($\gamma_{row},\gamma_{col},\gamma_{sparse}, ...$)\\ \hline
$\gamma$ & Gamma & Translates indices into offsets given a shape. \\ \hline
$\gamma^{'}$ & Gamma Inverse & Translates offsets into indices given a shape. \\ \hline
$\vec s \rshp \xi $ & Reshape& Changes the shape vector of an array,
 possibly affecting  \\
& &its dimensionality. Reshape depends on
layout($\gamma$). \\ \hline
$\pi \vec x$ & Pi & Returns a scalar and is equivalent to $\prod_{i=0}^{(\tau x)-1}x[i]$ \\ \hline
$\tau$& Tau & Returns the number of components in an array,($\tau \xi \equiv
\pi (\rho \xi) $) \\ \hline
$\xi_l \cat \xi_r $ & Catenate& Concatenates two arrays over their
primary axis. \\ \hline
$\xi_l f \xi_r $ & Point-wise  & A data parallel application of $f$ is performed \\
&Extension &  between all elements of the arrays.\\ \hline
$\sigma f \xi_r $ & Scalar Extension & $\sigma$ is used with
every component of $\xi_r$ in the data parallel \\
$\xi_l f \sigma$ & &application of $f$. \\ \hline
$\take$&Take& Returns a sub-array from the beginning or end of an array\\
& & based on its argument being positive or negative. \\ \hline
$\drop$ & Drop & The inverse of Take \\ \hline
$_{op} \mbox{red}$&Reduce& Reduce an array's dimension by one by applying \\
& & op over the primary axis of an array. \\ \hline
$\Phi$& Reverse& Reverses the components of an array. \\ \hline
$\Theta $& Rotate & Rotates, or shifts cyclically, components of an
array. \\ \hline
$\transpose$ & Transpose & Transposes the elements of an array based on \\
&&a given permutation vector \\ \hline
$\Omega$& Omega& Applies a unary or binary function to array argument(s) \\
&&given partitioning information. $\Omega$ is used to perform all operations \\&& (defined over the primary axis only) over all dimensions. \\ \hline
\end{tabular}
\caption{\label{moa}Summary of MoA Operations}
\end{figure}

@@

\subsection*{From Semantic to Operational Forms}

%Section with Psi  Correspondence Theorem and thinking in arrays *\

Consider the expression

\begin{eqnarray}
X & = & (2 \take ( \Phi A)) \times ( 1 \drop (\Phi A)) 
\end{eqnarray}
That is, take the first two planes after reversing A along the primary axis then multiply that array by the last two 
planes after reversing A. We want to concurrently perform both 5 by 4 multiplications. 
A is the array previously defined with shape $<3\;5\;4>$.
Suppose also that A is in shared memory and accessible to 2 processors.
First, we {\em Psi-Reduce} our expression to its Denotational Normal Form (DNF) or semantic 
normal form. This normal form can be used to prove the equivalence of  two array expressions.

Get shape: 
\begin{eqnarray}
\rho X & = & \rho (2 \take (\Phi A)) \nonumber \\
& = & 2 \cat ((1\drop (\rho ( \Phi A)))) \nonumber \\
&=& 2 \cat ((1 \drop (\rho  A))) \nonumber \\
&=& 2 \cat <\;5\;4> \nonumber \\
&=& <2\;4\;5>
\end{eqnarray}
Now that we have the {\em shape}, we can {Psi-Reduce}.
\[ \forall \;\;\; i,j,k \;\;\; \ni \;\;\; 0 \;\;\;\leq i,j,k < \;\;\;<2\;4\;5> \]
\[<i\;j\;k> \psi X \]
\begin{eqnarray}
& =&<i\;j\;k> \psi (2 \take ( \Phi A)) \times ( 1 \drop (\Phi A)) \nonumber \\
&=&<j\;k> \psi (<i> \psi (2 \take ( \Phi A)) \times ( 1 \drop (\Phi A)) )\nonumber \\
&= &<j\;k> \psi ((<i> \psi ( 2 \take (\Phi A))) \nonumber \\
&&\;\;\;\;\;\times ( <i> \psi ( 1 \drop (\Phi A)))) \nonumber \\
&= &<j\;k> \psi (( <i> \psi (\Phi A)) \times (<i+1> \psi (\Phi A))) \nonumber \\
&=&<j\;k> \psi ((<((\rho A)[0] - 1 - i)> \psi A ) \nonumber \\
&& \;\;\;\;\;\times  (<((\rho A)[0] - 1 -( i + 1))> \psi A ) )\nonumber \\
&=&<j\;k> \psi ((< 2-i> \psi A ) \times (<1-i>  \psi A) )
 \end{eqnarray}
\normalsize
This is the DNF, or semantic normal form. To proceed further we must know more about the data, i.e. layout, and architecture. 
\subsection{Applying the Psi Correspondence Theorem\cite{jen94}]
We now want to {\em transform} the DNF to a {\em Generic} Operational Normal Form (ONF) or way 
to {\em build the code}.
